New Biomass Pools Workflow

### 2023-08-23 NewBiomassPoolsWorkflow.Rmd
correction for benthos

calc_biomass_pool <- function(pooln, vol, biolprm){
  datalist <- list(pooln)

@@

@@

    pooln$atoutput <- with(pooln, vol * atoutput * bio_conv)

@@

biomass_pools <- calc_biomass_pool(pooln = testNpools,
                                   vol = vol,
                                   biolprm = biolprm)

@@

ggplot(txtbiopools, aes(x=time, y=atoutput)) +
  geom_line() +
  geom_point(data = aggbiopools, aes(x = time, y = totpool), colour = "blue") +
  theme_bw() +
  facet_wrap(~species, scales = "free_y")

@@

@@

```{r, warning=FALSE, fig.width=10}
mismatch <- txtbiopools |>
  dplyr::left_join(aggbiopools) |>
  dplyr::mutate(mismatch = atoutput - totpool,
                bad = ifelse(abs(mismatch) > 0.1*(atoutput), 1, 0))

@@

This is mismatch coded as 0 if less than 10% of txt output, 1 if more. Those exceeding are carrion, predatory benthos, detritivore benthos, king crab, benthic bacteria, large phytoplankton and sometimes small phytoplankton.

@@

```{r}
ggplot(txtbiopools |> filter(species %in% c("Predatory_ben", "Detrivore_ben")), 
       aes(x=time, y=atoutput)) +
  geom_line() +
  geom_point(data = aggbiopools |> filter(species %in% c("Predatory_ben", "Detrivore_ben")), 
             aes(x = time, y = totpool), colour = "blue") +
  theme_bw() +
  facet_wrap(~species, scales = "free_y")

>>>>

# we need area for the benthic inverts instead of volume
# THANK YOU JOE for this tip
area <- load_boxarea(dir = dir,
                     file_bgm = "Nordic02.bgm")

# we need a list of which group types to expand to area instead of volume
benthos <- c("SED_EP_FF", "SED_EP_OTHER", "SED_BAC", 
             "MOB_EP_OTHER", "LG_INF", "SM_INF", 
             "MICROPHYTBENTHOS", "SEAGRASS")

calc_biomass_pool <- function(pooln, vol, area, fgs, biolprm){
  datalist <- list(pooln)

@@

  # get fgs info
    # use grouptype column to allocate
  colnames(fgs) <- tolower(colnames(fgs))

  # check for GroupType or InvertType
  if (!("grouptype" %in% colnames(fgs) | "inverttype"%in% colnames(fgs))) {
    stop(paste("The columns GroupType or InvertType ars not in your functional\n",
               "groups file."))
  }

  # change inverttype to grouptype, contents should be the same
  if("inverttype" %in% names(fgs)) names(fgs)[names(fgs) == 'inverttype'] <- 'grouptype'

  fgs$grouptype <- tolower(fgs$grouptype)

  pooltype <- fgs |>
    dplyr::select(species=name, grouptype) |>
    dplyr::mutate(pooltype = dplyr::case_when((grouptype %in% c("lg_inf",
                                                                "sm_inf",
                                                                "sed_bact",
                                                                "sed_ep_ff",
                                                                "sed_ep_other",
                                                                "mobile_ep_other",
                                                                "coral",
                                                                "sponge")) ~ "benthos",
                                              TRUE ~ "nonbenth"))

@@

    pooln <- merge(pooln, area,
                   by = c("polygon"))

    pooln <- merge(pooln, pooltype,
                   by = c("species"))

    #pooln$atoutput <- with(pooln, vol * atoutput * bio_conv)

    pooln <- pooln |>
      dplyr::mutate(atoutput = dplyr::case_when(pooltype == "nonbenth" ~ vol * atoutput * bio_conv,
                                                pooltype == "benthos" ~ 
                                                  ifelse(layer==7, area * atoutput * bio_conv, 0)))

@@

biomass_pools <- calc_biomass_pool(pooln = testNpools,
                                   vol = vol,
                                   area = area,
                                   fgs = fgs,
                                   biolprm = biolprm)

@@

ggplot(txtbiopools, aes(x=time, y=atoutput)) +
  geom_line() +
  geom_point(data = aggbiopools, aes(x = time, y = totpool), colour = "blue", alpha=0.1) +
  theme_bw() +
  facet_wrap(~species, scales = "free_y")

@@

Joe pointed out that benthic groups should be expanded to area rather than volume. That correction has now been applied, and matches are better, but still diverge over time for some groups.

@@

```{r, warning=FALSE, fig.width=10}
mismatch <- txtbiopools |>
  dplyr::left_join(aggbiopools) |>
  dplyr::mutate(mismatch = atoutput - totpool,
                bad = ifelse(abs(mismatch) > 0.01*(atoutput), 1, 0))

@@

This is mismatch coded as 0 if less than 1% of txt output, 1 if more. Those exceeding are carrion, predatory benthos, detritivore benthos, king crab, benthic bacteria, large phytoplankton and sometimes small zooplankton.

@@

```{r}
ggplot(txtbiopools |> filter(species %in% c("Predatory_ben", "Detrivore_ben")), 
       aes(x=time, y=atoutput)) +
  geom_line() +
  geom_point(data = aggbiopools |> filter(species %in% c("Predatory_ben", "Detrivore_ben")), 
             aes(x = time, y = totpool), colour = "blue", alpha=0.1) +
  theme_bw() +
  facet_wrap(~species, scales = "free_y")
